---
title: "In-Class Exercise 03"
author: "Seah Chi Hao"
date: 2024-09-02
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
  output: true
  warning: false

---

## 1.0 Installing maptools

```{r}
#| echo: false
pacman::p_load(sf, raster, spatstat, tmap, tidyverse)
```

**maptools** is retired and binary is removed from CRAN. However, we can download from [Posit Public Package Manager](https://packagemanager.posit.co) snapshots by using the code chunk below.
```{r}
#| eval: false #so that this chunk won't keep running (add this when there is adhoc installation)
#After the installation is complete, it is important to edit the code chunk to avoid maptools being downloaded and installed each time.
#By adding "#| eval: false"
install.packages("maptools",
                 repos = "https://packagemanager.posit.co/cran/2023-10-13")
```
> Note: Need to know which version of package you want. Ususally use this method when it's not available in the achive of CRAN.

> Note: Try not to use maptools since it is already deprecate.

## 2.0 Creating costal outline

### 2.1 Working with st_union()
The code chunk below, `st_union()` is used to derive the costal outline sf tibble data.frame.

```{r}
mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")
sg_sf <- mpsz_sf %>%
  st_union()
```

## 3.0 Read Myanmar data

```{r}
acled_sf <- read_csv("data/ACLED_Myanmar.csv") %>%
  st_as_sf(coords = c(
    "longitude", "latitude"),
    crs=4326) %>%
  st_transform(crs = 32647) %>%
  mutate(event_date = dmy(event_date))
```